Cluster Difference Imaging Photometric Survey. IV. TOI 1937Ab. Youngest Hot Jupiter, or Just Tidally Spun Up?

### 2021-07-16 cluster_paper/ms.tex
ngc2516: AJ accepted

%% Reintroduced the \received and \accepted commands from AASTeX v5.2.
%% Add "Submitted to " argument.
\received{April 29, 2021}
\revised{July 7, 2021}
\accepted{---}
\submitjournal{AAS journals.}
\shorttitle{The Halo of NGC\,2516}

@@

\added{To assess whether the stars in the cluster's halo are the same
ages as stars in the cluster's core, we apply three different age-dating
techniques: isochrones, gyrochrones, and lithium depletion
\citep[see][for reviews]{soderblom_ages_2010,soderblom_ages_2014}.
Gyrochronology has now been empirically calibrated for FGK stars with
ages spanning $\approx$10\,Myr to $\approx$2.7\,Gyr \citep[{\it
e.g.},][]{rebull_usco_2018,curtis_temporary_2019,curtis_rup147_2020}.
Lithium equivalent widths can be used for relative age-dating of dwarf
stars over similar age ranges, though for a narrower range of spectral
types \citep{sestito_2005}.  One complication when deriving relative
lithium ages of the K dwarfs comes from the correlation between
lithium abundance and stellar rotation rate, which has recently been
reviewed by \citet{bouvier_lithium-rotation_2020}, and is a subject
that we explore in the latter portions of our analysis.  }

@@

Fitting a model to substantiate the claim that \cn\ is
gyrochronologically older than the Pleiades \citep[{\it e.g.},][]{mamajek_improved_2008,angus_toward_2019,spada_competing_2020}
is, unfortunately, an exercise in tautology.  Gyrochronological models
are empirically calibrated against the Pleiades and Praesepe.  At
$\bpmrpo=1.35$, we observe rotation periods in \cn\ ranging from 7 to
10 days.  The 7 day rotation periods are consistent with the Pleiades;
the 10 day rotation periods are not.  Fitting formulae from the
previously cited studies give ages for a star with $\{\bpmrpo=1.35,
(\mathrm{B}-\mathrm{V})_0=1.10, P=7\,{\rm day}\}$ of 204 Myr, 316 Myr,
and 107 Myr respectively.  Given a Pleiades age of $125\pm20$\,Myr (an
average of MSTO and rotation-corrected lithium-depletion results; see
{\it e.g.} \citealt{stauffer_keck_1998},
\citealt{soderblom_ages_2014}, and \citealt{cummings_2018}), only the
\citet{spada_competing_2020} model has an absolute scale that seems to
be well-calibrated at these ages and colors.  For an 8 day rotation
period of a star with the same color, this model quotes a 193 Myr age;
at 9 days, 388 Myr.  Given this degree of model uncertainty, we prefer
the relative statement that \cn\ appears slightly gyrochronologically
older than the Pleiades, and much younger than Praesepe.  A
gyrochronological age estimate for \cn\ between 150 and 200\,Myr would
therefore appear reasonable.  Much older, and the issue of why the F
and G dwarfs do not spin down becomes pressing.  We caution however
that the spindown rate of FGK stars between 100 and 300 Myr has not
yet been empirically calibrated, and if ``stalling'' were to occur, it
could affect our age assessment \citep[see][]{curtis_rup147_2020}.

@@

We performed the same light curve stitching and period-search
procedure discussed in the text on the field stars.
\replaced{Imposing}{Within $0.4<\bpmrpo<0.8$}, the same requirements
for crowding resulted in \ncompstardenominator\ \added{field} stars
for which rotation periods could have been detected.  Imposing the
same Lomb Scargle power cutoff (${\rm LSP}>0.08$) and \replaced{period
upper limit ($P<15\,{\rm days}$)}{color-period cutoff as that used to
define Set $\mathcal{B}$} yielded \ncompstarnumerator\ period
detections (\ncompfrac).  Within the same \replaced{brightness
cutoff}{color range}, \nautovscompstarnumerator\ of
\nautovscompstardenominator\ \replaced{cluster stars}{kinematically
identified candidate cluster members} yielded period detections
(\nautofrac).  \replaced{While the detection fractions are only mildly different
(likely due to the brightness cutoff), the }{The }period--color distributions
are \replaced{very}{also quite} different\added{: the candidate
cluster members show an overdensity coincident with the expected
gyrochronal age, while the field stars show a much broader rotation
period distribution} (Figure~\ref{fig:compstar}).  This confirms our
expectation that our period measurement procedure is not biased in any
way that could produce the features we observe in \cn.

>>>>

%% Reintroduced the \received and \accepted commands from AASTeX v5.2.
%% Add "Submitted to " argument.
\received{April 29, 2021}
\revised{July 7, 2021}
\accepted{July 16, 2021}
\submitjournal{AAS Journals.}
\shorttitle{The Halo of NGC\,2516}

@@

\added{To assess whether the stars in the cluster's halo are the same
ages as stars in the cluster's core, we apply three different age-dating
techniques: isochrones, gyrochrones, and lithium depletion
\citep[see][for reviews]{soderblom_ages_2010,soderblom_ages_2014}.
Gyrochronology has now been empirically calibrated for FGK stars with
ages spanning $\approx$10\,Myr to $\approx$4\,Gyr \citep[{\it
e.g.},][]{rebull_usco_2018,curtis_temporary_2019,curtis_rup147_2020,barnes_rotation_2016}.
Lithium equivalent widths can be used for relative age-dating of dwarf
stars over similar age ranges, though for a narrower range of spectral
types \citep{sestito_2005}.  One complication when deriving relative
lithium ages of the K dwarfs comes from the correlation between
lithium abundance and stellar rotation rate, which has recently been
reviewed by \citet{bouvier_lithium-rotation_2020}, and is a subject
that we explore in the latter portions of our analysis.  }

@@

Fitting a model to substantiate the claim that \cn\ is
gyrochronologically older than the Pleiades \citep[{\it e.g.},][]{mamajek_improved_2008,angus_toward_2019,spada_competing_2020}
is, unfortunately, an exercise in tautology.  Gyrochronological models
are empirically calibrated against the Pleiades and Praesepe.  At
$\bpmrpo=1.35$, we observe rotation periods in \cn\ ranging from 7 to
10 days.  The 7 day rotation periods are consistent with the Pleiades;
the 10 day rotation periods are not.  Fitting formulae from the
previously cited studies give ages for a star with $\{\bpmrpo=1.35,
(\mathrm{B}-\mathrm{V})_0=1.10, P=7\,{\rm day}\}$ of 204 Myr, 316 Myr,
and 107 Myr respectively.  Given a Pleiades age of $125\pm20$\,Myr (an
average of MSTO and rotation-corrected lithium-depletion results; see
{\it e.g.} \citealt{stauffer_keck_1998},
\citealt{soderblom_ages_2014}, and \citealt{cummings_2018}), only the
\citet{spada_competing_2020} model has an absolute scale that seems to
be well-calibrated at these ages and colors.  For an 8-day rotation
period of a star with the same color, this model quotes a 193 Myr age;
at 9 days, 388 Myr.  Given this degree of model uncertainty, we prefer
the relative statement that \cn\ appears slightly gyrochronologically
older than the Pleiades, and much younger than Praesepe.  A
gyrochronological age estimate for \cn\ between 150 and 200\,Myr would
therefore appear reasonable.  Much older, and the issue of why the F
and G dwarfs do not spin down becomes pressing.  We caution however
that the spindown rate of FGK stars between 100 and 300 Myr has not
yet been empirically calibrated, and if ``stalling'' were to occur, it
could affect our age assessment \citep[see][]{curtis_rup147_2020}.

@@

We performed the same light curve stitching and period-search
procedure discussed in the text on the field stars.
\replaced{Imposing}{Within $0.4<\bpmrpo<0.8$}, the same requirements
for crowding resulted in \ncompstardenominator\ \added{field} stars
for which rotation periods could have been detected.  Imposing the
same Lomb-Scargle power cutoff (${\rm LSP}>0.08$) and \replaced{period
upper limit ($P<15\,{\rm days}$)}{color-period cutoff as that used to
define Set $\mathcal{B}$} yielded \ncompstarnumerator\ period
detections (\ncompfrac).  Within the same \replaced{brightness
cutoff}{color range}, \nautovscompstarnumerator\ of
\nautovscompstardenominator\ \replaced{cluster stars}{kinematically
identified candidate cluster members} yielded period detections
(\nautofrac).  \replaced{While the detection fractions are only mildly different
(likely due to the brightness cutoff), the }{The }period--color distributions
are \replaced{very}{also quite} different\added{: the candidate
cluster members show an overdensity coincident with the expected
gyrochronal age, while the field stars show a much broader rotation
period distribution} (Figure~\ref{fig:compstar}).  This confirms our
expectation that our period measurement procedure is not biased in any
way that could produce the features we observe in \cn.

### 2021-07-16 planet_paper/ms.tex
ngc2516: AJ accepted

>>>>

- Get a spectrum of the secondary?
  Andrew Mann suggests this is possible.
     The primary is G=13.0, secondary G=17.6.  The age test would be
     based on Halpha.  
     ``If there's no halpha you can be sure it's old, but if there is
     some it's kinda inconclusive''
     ``10-40\% of M3s are inactive, assuming that the SpT is right ''
     ``get the M dwarf spectrum in good seeing and confirm age, or
     don't bother''
